Hypercoercions and a Framework for \\ Equivalence of Cast Calculi

### 2020-01-05 paper.tex
fixed fig 5
fixed statement about common-case coercions

\begin{figure}
  \fbox{$\mathit{applyCast}(v,c) = r$}
  \[
  \begin{array}{rclr}
    \mathit{applyCast}(v, \cOOcast{\star}{l}{\star} ) &=& v \\
    \mathit{applyCast}(v, \cOOcast{P}{l}{\star}) &=&
        v \langle \cOOcast{P}{l}{I} \rangle
          \langle \cOOcast{I}{l}{\star} \rangle
        & \text{if } I \sim P, I \neq P \\  
    \mathit{applyCast}(v, \cOOcast{\star}{l}{P}) &=&          
        v \langle \cOOcast{\star}{l}{I} \rangle
          \langle \cOOcast{I}{l}{P} \rangle
        & \text{if } I \sim P, I \neq P \\  
  \mathit{applyCast}(v \langle \cOOcast{I}{l}{\star} \rangle , \cOOcast{\star}{l}{I}) &=& v \\
  \mathit{applyCast}(v \langle \cOOcast{I_1}{l}{\star} \rangle , \cOOcast{\star}{l}{I_2}) &=& \rOOfail{l} & \text{if } I_1 \neq I_2 \\
  \mathit{applyCast}(v, \cOOcast{P_1}{l}{P_2}) &=&
     v \langle \cOOcast{P_1}{l}{P_2} \rangle & \text{if } P_1 \smile P_2
  \end{array}
  \]

@@

Hypercoercions enable a bit-level representation that is particularly
compact for the most common-case coercions. The most frequent
coercions are injections and projections to base types (here
$\POOunit$ is the only base type, but in a real language the base
types would include integers, Booleans, etc.) and identity coercions.
Values of base types are often stored in CPU registers, so it would be
nice for coercions on base types to also fit in registers, that is, in
64 bits, so that applying a coercion to a value would not require
access to main memory, which is an order of magnitude slower than
accessing registers on a modern CPU.
%
Coercions involving non-base types, such as function types,
may be arbitrarily large, so in those cases, the hypercoercion
representation can be a pointer to a heap-allocated structure.  Here
is a sketch for the bit-level representation for hypercoercions.

>>>>

\begin{figure}
  \fbox{$\mathit{applyCast}(v,c) = r$}
  \[
  \begin{array}{rclr}
    \mathit{applyCast}(v, \cOOcast{\star}{l}{\star} ) &=& \rOOsucc{v} \\
    \mathit{applyCast}(v, \cOOcast{P}{l}{\star}) &=&
        \rOOsucc{v \langle \cOOcast{P}{l}{I} \rangle
                   \langle \cOOcast{I}{l}{\star} \rangle}
        & \text{if } I \sim P, I \neq P \\  
    \mathit{applyCast}(v, \cOOcast{\star}{l}{P}) &=&          
        \rOOsucc{v \langle \cOOcast{\star}{l}{I} \rangle
                   \langle \cOOcast{I}{l}{P} \rangle}
        & \text{if } I \sim P, I \neq P \\  
  \mathit{applyCast}(v \langle \cOOcast{I}{l}{\star} \rangle , \cOOcast{\star}{l}{I}) &=& \rOOsucc{v} \\
  \mathit{applyCast}(v \langle \cOOcast{I_1}{l}{\star} \rangle , \cOOcast{\star}{l}{I_2}) &=& \rOOfail{l} & \text{if } I_1 \neq I_2 \\
  \mathit{applyCast}(v, \cOOcast{P_1}{l}{P_2}) &=&
           \rOOsucc{v \langle \cOOcast{P_1}{l}{P_2} \rangle}
          & \text{if } P_1 \smile P_2
  \end{array}
  \]

@@

Hypercoercions enable a bit-level representation that is particularly
compact for identity coercions and coercions that project from base
types or inject to base types.  (Here $\POOunit$ is the only base
type, but in a real language the base types would include integers,
Booleans, etc.)  We conjecture that such coercions occur more
frequently than the more complex coercions (e.g. between function
types), especially because such coercions appear in the leaves of
complex coercions.

Futhermore, values of base types are often stored in CPU registers, so
it would be nice for coercions on base types to also fit in registers,
that is, in 64 bits, so that applying a coercion to a value of base
type would not require access to main memory, which is an order of
magnitude slower than accessing registers on a modern CPU.
%
Coercions involving non-base types, such as function types, may be
arbitrarily deep, so in those cases, the hypercoercion representation
has to be a pointer to a heap-allocated structure.

Here is a sketch for the bit-level representation for hypercoercions.
